---
title: "R Notebook"
output: html_notebook
---
```{r}
library(lme4)
library(readr)
big_lm <- read_csv("C:/Users/kayla/Downloads/biglinreg.csv")

library(ggplot2)

ggplt <- ggplot(data = big_lm, aes(x=average_plantcover,y=seedbank_density,shape=season), labs(main = "Seedbank Density Increases with BFT Cover Across All Sites and Between Seasons",
x = "BFT Aboveground Cover (%)",
y = "Seedbank Density (seeds/m2)"))+
         geom_point()+
         theme_classic()

ggplt

# Plotting multiple Regression Lines
ggplt+geom_smooth(method=lm,se=FALSE,fullrange=TRUE,
                  aes(color=season))
``` 
```{r}
library(lme4)
library(readr)
big_lm <- read_csv("C:/Users/kayla/Downloads/biglinreg.csv")

# Fit the Generalized Linear Mixed Model
# The formula syntax is: response ~ fixed_effects + (random_effects | grouping_factor)
# The 'family = binomial' specifies a logistic regression with a logit link function
model_big <- lm(
  seedbank_density ~ average_plantcover + season + site,
  data = big_lm,
)

# View the model summary
summary(model_big)

library(car)
type3_anovabig <- Anova(model_big, type = "III")
print(type3_anovabig)
```
```{r}
library(lme4)
library(readr)
big_lm <- read_csv("C:/Users/kayla/Downloads/biglinreg.csv")

# Fit the Generalized Linear Mixed Model
# The formula syntax is: response ~ fixed_effects + (random_effects | grouping_factor)
# The 'family = binomial' specifies a logistic regression with a logit link function
model_big2 <- lm(
  average_plantcover ~ season + site,
  data = big_lm,
)

# View the model summary
summary(model_big2)

library(car)
type3_anovabig2 <- Anova(model_big2, type = "III")
print(type3_anovabig2)
```

```{r}
library(readr)
Linreg_spring26 <- read_csv("C:/Users/kayla/Downloads/Spring26data/spring26linreg.csv")

coefficient_s26 <- cor.test(Linreg_spring26$average_plantcover, Linreg_spring26$seedbank_density)

coefficient_s26$estimate

model_s26 <- lm(seedbank_density ~ average_plantcover, data = Linreg_spring26)

summary(model_s26)
```


```{r}
library(readr)
Linreg_spring26 <- read_csv("C:/Users/kayla/Downloads/Spring26data/spring26linreg.csv")

x <- Linreg_spring26$average_plantcover
y <- Linreg_spring26$seedbank_density
# Plot with main and axis titles
# Change point shape (pch = 19) and remove frame.
p<- plot(x, y, main = "Site Wide BFT Cover vs Seedbank Density Spring 2026",
    xlab = "BFT Cover (%)", ylab = "Seedbank Density (seeds/m2)",
     pch = 19, frame = FALSE)
abline(lm(y ~ -1 + x, data = Linreg_spring26), col = "blue")
```
```{r}
library(readr)
Linregsample_spring26 <- read_csv("C:/Users/kayla/Downloads/Spring26data/spring26samplelevellinreg.csv")

x <- Linregsample_spring26$BFT_percent
y <- Linregsample_spring26$Seedbank_Density
# Plot with main and axis titles
# Change point shape (pch = 19) and remove frame.
p<- plot(x, y, main = "Sample Level BFT Cover vs Seedbank Density Spring 2026",
    xlab = "BFT Cover (%)", ylab = "Seedbank Density (seeds/m2)",
     pch = 19, frame = FALSE)
abline(lm(y ~ -1 + x, data = Linregsample_spring26), col = "blue")
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
```{r}
library(readr)
Linregsample_spring26 <- read_csv("C:/Users/kayla/Downloads/Spring26data/spring26samplelevellinreg.csv")

coefficient_ss26 <- cor.test(Linregsample_spring26$BFT_percent, Linregsample_spring26$Seedbank_Density)

coefficient_ss26$estimate

model_ss26 <- lm(Seedbank_Density ~ BFT_percent, data =Linregsample_spring26)

summary(model_ss26)
```
```{r}
library(readr)
Springvspring <- read_csv("C:/Users/kayla/Downloads/Spring25v26boxplots.csv")
# library
library(ggplot2)
 
ggplot(data = Springvspring, aes(x=Invasion, y=Seedbank, fill=Season)) + 
    geom_boxplot(aes(x=Invasion, y=Seedbank, fill=Season)) + scale_color_hue() + theme_light()
```
```{r}
library(readr)
sp26boxplot <- read_csv("C:/Users/kayla/Downloads/spring26boxplot.csv")


boxplot(sp26boxplot$No_invasion_seedbank_density, sp26boxplot$Low_invasion_seedbank_density, sp26boxplot$High_invasion_seedbank_density,
main = "Seedbank Density Across Invasion Levels Spring 2026",
xlab = "Invasion Level",
ylab = "Seedbank Density (seeds/m2)",
names = c("No Invasion", "Low Invasion","High Invasion"),
col = c("yellow", "green", "blue"),
border = "black",
Vertical = TRUE,
notch = FALSE
)
```
```{r}

library(readr)
spboxplot <- read_csv("C:/Users/kayla/Downloads/Spring26data/boxplot_sp2526.csv")


boxplot(spboxplot$No_invasion_seedbank_density, spboxplot$Low_invasion_seedbank_density, spboxplot$High_invasion_seedbank_density,
main = "BFT Seedbank Density Across Invasion Levels Spring",
xlab = "Invasion Level",
ylab = "Seedbank Density (seeds/m2)",
names = c("No Invasion", "Low Invasion","High Invasion"),
col = c("yellow", "green", "blue"),
border = "black",
Vertical = TRUE,
notch = FALSE
)

```
```{r}

library(readr)
violin_allspring <- read_csv("C:/Users/kayla/Downloads/Spring26data/violin_springalldata.csv")

library(ggplot2)

library(scales)
# Basic violin plot
sp_violin <- ggplot(data = violin_allspring, aes(x = Invasion, y = Seedbank_Density)) +
   geom_violin(fill = "white", alpha = 0.5, adjust = 2, scale = "width") + geom_jitter(width = 0.08, alpha = 0.5, color = "#3a86d4")


sp_violin

```

```{r}
library(readr)
sp26boxplot <- read_csv("C:/Users/kayla/Downloads/spring26boxplot.csv")


boxplot(sp26boxplot$Low_invasion_seedbank_density,
main = "Low Invasion Seedbank Density",
xlab = "Invasion Level",
ylab = "Seedbank Density (seeds/m^2)",
names = "Low Invasion",
col = "blue",
border = "green",
Vertical = TRUE,
notch = TRUE
)
```
```{r}
library(lme4)
library(readr)
spvs_lm <- read_csv("C:/Users/kayla/Downloads/Spring26data/lme_sp25v26.csv")

# Fit the Generalized Linear Mixed Model
# The formula syntax is: response ~ fixed_effects + (random_effects | grouping_factor)
# The 'family = binomial' specifies a logistic regression with a logit link function
model_spvs <- lm(
 seedbank_density ~ average_plantcover + year,
  data = spvs_lm,
)

# View the model summary
summary(model_spvs)

library(car)
type3anova_spvs <- Anova(model_spvs, type = "III")
print(type3anova_spvs)
```
```{r}
library(lme4)
library(readr)
spvsinvasion_lm <- read_csv("C:/Users/kayla/Downloads/Spring26data/lme_sp25v26.csv")

# Fit the Generalized Linear Mixed Model
# The formula syntax is: response ~ fixed_effects + (random_effects | grouping_factor)
# The 'family = binomial' specifies a logistic regression with a logit link function
model_spvsinvasion <- lm(
 seedbank_density ~ invasion_level + year,
  data = spvsinvasion_lm,
)

# View the model summary
summary(model_spvsinvasion)

library(car)
type3anova_spvsinvasion <- Anova(model_spvsinvasion, type = "III")
print(type3anova_spvsinvasion)
```
```{r}
#spring 25 vs 26 does not have a significant difference overall between seasons, so I can combine the data into one 
library(readr)
lowttest_spvs <- read_csv("C:/Users/kayla/Downloads/Spring26data/sp25v26_lowttest.csv")
t.test(lowttest_spvs$low_sp25, lowttest_spvs$low_sp26, alternative = "greater")
wilcox.test(lowttest_spvs$low_sp25, lowttest_spvs$low_sp26, alternative = "greater")

#low invasion sites vary from one another between spring 25 and 26
```
```{r}
library(lme4)
library(readr)
spvshi_lm <- read_csv("C:/Users/kayla/Downloads/Spring26data/sp25v26_hisites.csv")

# Fit the Generalized Linear Mixed Model
# The formula syntax is: response ~ fixed_effects + (random_effects | grouping_factor)
# The 'family = binomial' specifies a logistic regression with a logit link function
model_spvshi <- lm(
 seedbank_density ~ site + year,
  data = spvshi_lm,
)

# View the model summary
summary(model_spvshi)

library(car)
type3anova_spvshi <- Anova(model_spvshi, type = "III")
print(type3anova_spvshi)

#high invasion plot by plot did not change significantly between spring 25 vs 26
```


